\subsection{Constructions with Compactness~(2.3.2)}

\begin{proposition}
	The unit interval~$[0, 1]$ is compact.
\end{proposition}

\begin{proof}
	Let~$\cover{U}$ be an open cover of~$[0, 1]$.
	Let
	\[
		F ≔ \{ u ∈ [0, 1] \suchthat \text{$[0, u]$ is covered by a finite subset of~$\cover{U}$} \} \,,
	\]

	There exists for every point~$x$ in~$X$ a set~$U$ belonging to~$\cover{U}$ with~$x ∈ U$.
	For every point~$y ∈ U$, the initial segment~$[0, x]$ is covered by a finite subset of~$\cover{U}$ if and only if the segment~$[0, y]$ is covered by a finite subset of~$U$.
	Therefore, if~$x ∈ F$, then~$U ⊆ F$, and if otherwise~$x ∈ [0, 1] ∖ F$, then~$U ⊆ [0, 1] ∖ F$.
	This shows that both~$F$ and~$[0, 1] ∖ F$ are open.

	It follows that either~$F = ∅$ or~$F = [0, 1]$, because~$[0, 1]$ is connected.
	We know that~$0$ is contained in~$F$, so~$F$ is nonempty.
	Therefore,~$F = [0, 1]$.
\end{proof}



\subsubsection{Subspaces of Compact Spaces}

The interval~$[0, 1]$ is a compact Hausdorff space, and~$(0, 1)$ is a non-closed subspace of~$[0, 1]$.
The subspace~$(0, 1)$ is therefore not compact.



\subsubsection{Quotients of Compact Spaces}

Let~$S$ be a quotient space of a compact topological space~$X$.
The canonical quotient map from~$X$ to~$S$ is surjective and continuous, so it follows from Theorem~2.15 that~$S$ is compact.



\subsubsection{Coproducts of Compact Spaces}

The book simply claims that \enquote{coproducts of compact spaces are certainly not compact}.
However, this isn’t quite true:
coproducts of \emph{finitely} many compact spaces are again compact, as we will now show:

Let~$X_1, \dotsc, X_n$ be finitely many compact topological spaces and let
\[
	X ≔ X_1 ⨿ \dotsb ⨿ X_n
\]
be their coproduct.
Let~$\cover{U}$ be an open cover of~$X$.
Each set~$U$ of~$\cover{U}$ is of the form
\[
	U = U_1 ⨿ \dotsb ⨿ U_n
\]
for open subsets~$U_i$ of~$X_i$.
That~$\cover{U}$ is an open cover of~$X$ is equivalent to
\[
	\{ U_i \suchthat U ∈ \cover{U} \}
\]
being an open cover of~$X_i$ for every~$i = 1, \dotsc, n$.
It follows for every~$i = 1, \dotsc, n$ that there exists a finite subset~$\cover{U}_i$ of~$\cover{U}$ such that~$\{ U_i \suchthat U ∈ \cover{U}_i \}$ is an open cover of~$X_i$.
It follows that~$⋃_{i = 1}^n \cover{U}_i$ is a finite subset of~$\cover{U}$ that still covers~$X$.

However, the coproduct of infinitely many nonempty compact spaces is never again compact:
if~$(X_α)_{α ∈ A}$ is an infinite family of nonempty compact spaces, then~$\{ X_α \suchthat α ∈ A \}$ is an open cover of~$∐_{α ∈ A} X_α$ that does not admit any proper subcover, and hence in particular no finite subcover.
