\subsection{Exercise~2.6}
\label{exercise 2.6}

This is nothing special to the connected components:

\begin{proposition}
	Let~$X$ be a topological space and let~$(X_α)_{α ∈ A}$ be a family of subsets with~$X = ∐_{α ∈ A} X_α$ as sets.
	Let~$∼$ be the equivalence relation on~$X$ that identifies each~$X_α$ to a single point.
	The following conditions are equivalent:
	\begin{equivalenceslist}

		\item
			\label{is coproduct}
			The topological space~$X$ is the coproduct of the spaces~$X_α$.

		\item
			\label{each subspace is open}
			Each~$X_α$ is open in~$X$.

		\item
			\label{quotient space is discrete}
			The quotient space~$X / {∼}$ is discrete.

	\end{equivalenceslist}
\end{proposition}

\begin{proof}
	We show that the first two conditions are equivalent, and then that the last two conditions are equivalent.
	\begin{implicationslist}

	\item[\ref{is coproduct}~$\implies$~\ref{each subspace is open}]
		Whenever we have an open subset~$U_α$ of~$X_α$ for every index~$α ∈ A$, the set~$∐_{α ∈ A} U_α$ is open in~$X$.
		For every index~$β$ we may choose these sets~$U_α$ as~$U_β = X_β$ and~$U_α = ∅$ for every~$α ≠ β$ to find that~$∐_{α ∈ A} U_α = X_β$ is open in~$X$.

	\item[\ref{each subspace is open}~$\implies$~\ref{is coproduct}]
		We need to show that an arbitrary subset~$U$ of~$X$ is open in~$X$ if and only if it is of the form~$U = ∐_{α ∈ A} U_α$ for open subsets~$U_α$ of~$X_α$.

		Suppose that~$U$ is open in~$X$.
		Then each intersection~$U_α ≔ U ∩ X_α$ is again open in~$X$, and therefore also open in~$X_α$, and we have~$U = ⋃_{α ∈ A} U_α = ∐_{α ∈ A} U_α$.

		Suppose conversely that the set~$U = ∐_{α ∈ A} U_α$ for open subsets~$U_α$ of~$X_α$.
		Each~$U_α$ is also open in~$X$, because~$U_α$ is open in~$X_α$ and~$X_α$ is open in~$X$.
		It follows that~$U = ∐_{α ∈ A} U_α = ⋃_{α ∈ A} U_α$ is again open in~$X$.

	\item[\ref{each subspace is open}~$\implies$~\ref{quotient space is discrete}]
		Each~$X_α$ is a saturated open subset of~$X$.
		This is equivalent to saying that every point of~$X / {∼}$ is open, i.e., that every singleton subset of~$X / {∼}$ is open.
		This means that~$X / {∼}$ is discrete.

	\item[\ref{quotient space is discrete}~$\implies~$\ref{each subspace is open}]
		Let~$α$ be an index.
		If the set~$X_α$ is empty then it is in particular open in~$X$.
		If~$X_α$ is nonempty, then its image in~$X / {∼}$ is a single point~$y$.
		The set~$X_α$ is then the preimage of the singleton set~$\{ y \}$, which is open in~$X / {∼}$, and therefore open in~$X$ by the continuity of the quotient map.
	\qedhere

	\end{implicationslist}
\end{proof}
