\subsection{Exercise~2.23}

We generalize Theorem~2.14 (page~49) as follows:

Let~$A$ be an arbitrary index set of cardinality at least two.
Let~$X^A$ be the product~$∏_{α ∈ A} X$.
We show that the following conditions on~$X$ are equivalent:
\begin{equivalenceslist*}

	\item
		The space~$X$ is a Hausdorff space.

	\item
		The diagonal set~$Δ_{X, A} = \{ (x_α)_α ∈ X^A \suchthat \text{$x_α = x_β$ for all~$α, β ∈ A$} \}$ is closed in~$X × X$.

	\item
		The diagonal map~$Δ \colon X \to X^A$,~$x \mapsto (x)_α$ is closed.

\end{equivalenceslist*}
We first show that the first two conditions are equivalent, and then we show that the last two conditions are equivalent.

Suppose first that~$X$ is a Hausdorff space.
We then consider for every two indices~$β, γ ∈ A$ the projection map
\[
	π_{β γ} \colon X^A \to X × X \,, \quad (x_α)_α \mapsto (x_β, x_γ) \,,
\]
which is continuous.
We know from Theorem~2.14 that the diagonal set~$Δ_X$ in~$X × X$ is closed.
It follows that
\[
	Δ_{X, A} = ⋂_{β, γ ∈ A} π_{β γ}^{-1} Δ_X
\]
is closed in~$X^A$.

Suppose now that the diagonal set~$Δ_X$ is closed in~$X^A$.
We choose an index~$β ∈ A$ and consider the map
\[
	f
	\colon
	X × X
	\to
	X^A \,,
	\quad
	(x, y)
	\mapsto
	(x_α)_α
	\quad\text{with}\quad
	x_α
	=
	\begin{cases*}
		x & if~$α = β$, \\
		y & otherwise.
	\end{cases*}
\]
This map is continuous, and we have~$Δ_X = f^{-1} Δ_{X, A}$ since~$A$ contains at least one other index apart from~$β$.
It follows that~$Δ_X$ is closed in~$X × X$.
According to Theorem~2.14, this shows that~$X$ is a Hausdorff space.

For the second equivalence, we note that the two maps
\[
	φ \colon X \to Δ_{X, A} \,, \quad x \mapsto (x)_α
\]
and
\[
	ψ \colon Δ_{X, A} \to X \,, \quad (x)_α \mapsto x
\]
are mutually inverse and continuous, whence~$φ$ is a homeomorphism.
The map~$φ$ is also the corestriction of~$Δ$ to its image.
We have therefore the chain of equivalences
\begin{align*}
	{}&
	\text{the map~$Δ$ is closed} \\
	\iff{}&
	\left\{
	\begin{tabular}{l}
		the image of~$Δ$ is closed in~$X^A$, \\
		and the corestricion of~$Δ$ to its image is closed
	\end{tabular}
	\right. \\
	\iff{}&
	\text{the set~$Δ X$ is closed and the map~$φ$ is closed} \\
	\iff{}&
	\text{the set~$Δ_{X, A}$ is closed} \,,
\end{align*}
since the homeomorphism~$φ$ is automatically closed.
