\subsection{Exercise~1.6}
\label{exercise 1.6}

\begin{definition}
	A subset~$D$ of a topological space~$X$ is \defemph{dense} if it intersects every nonempty open subset of~$X$ non-trivially. That is, for every nonempty open subset~$U$ of~$X$, the intersection~$D ∩ U$ is again nonempty.
\end{definition}

\begin{definition}
	A topological space is \defemph{separable} if it admits a countable dense subset.
\end{definition}

Separability is a homeomorphism invariant:
if two topological spaces~$X$ and~$Y$ are homeomorphic, then~$X$ is separable if and only if~$Y$ is separable.

We show in the following that~$ℓ_p$ is separable for every~$1 ≤ p < ∞$, but that~$ℓ_∞$ is not separable.
To this end, we make the following two observations.

\begin{lemma}
	\label{characterization of dense subsets of metric spaces}
	A subset~$D$ of a metric space~$(X, d)$ is dense if and only if there exists for every~$x ∈ X$ and every~$ε > 0$ some~$y ∈ D$ with~$d(x, y) < ε$.
	\qed
\end{lemma}

\begin{proposition}
	\label{criterion for non-separable in metric spaces}
	Let~$X$ be a metric space.
	Suppose there exists an uncountable subset~$A$ of~$X$ and a constant~$C > 0$ such that~$d(a, b) > C$ for any two distinct points~$a$ and~$b$ in~$A$.
	Then~$X$ is not separable.
\end{proposition}

\begin{proof}
	For every point~$a$ in~$A$ let~$U_a ≔ \ball(a, C / 2)$.
	The sets~$U_a$ are open in~$X$ and pairwise disjoint.
	
	Let~$D$ be a dense subset of~$X$, and for every point~$a$ in~$A$ let~$D_a = D ∩ U_a$.
	Each set~$D_a$ must be nonempty because~$D$ is dense and~$U_a$ is open.
	The sets~$D_a$ are also pairwise disjoint because the sets~$U_a$ are pairwise disjoint.

	The sets~$D_a$ thus form a partition for the subset~$D' ≔ ⋃_{a ∈ A} D_a$ of~$D$.
	We have therefore a surjective map from~$D'$ to~$A$, assigning to each point~$x$ of~$D'$ the unique point~$a$ in~$A$ with~$x ∈ D_a$.
	It follows that~$\card{D} ≥ \card{D'} ≥ \card{A}$.
	As~$A$ is uncountable,~$D$ is also uncountable.
\end{proof}

To show that~$ℓ_p$ is separable for~$1 ≤ p < ∞$ we consider its subset~$D$ of all sequences~$(x_n)_n$ such that all~$x_n$ are rational and only finitely many~$x_n$ are nonzero.
If we think about the space~$ℓ_p$ as a subset of the countable product~$ℝ^ℕ = ℝ × ℝ × \dotsb$, then the set~$D$ is given by
\[
	D = ⋃_{n ≥ 0} \underbrace{ℚ × \dotsb × ℚ}_{\text{$n$ times}} × \{0\} × \{0\} × \dotsb
\]
We see from this description that~$D$ is a countable union of countable sets, and therefore again countable.

Let~$x$ be a sequence in~$\ell_p$ and let~$ε > 0$.
We have for every~$m ≥ 0$ a truncated sequence~$x^{≤ m}$ given by
\[
	x^{≤ m}_n
	=
	\begin{cases*}
		x_n & for~$n ≤ m$, \\
		0   & for~$n > m$.
	\end{cases*}
\]
This is again a sequence in~$ℓ_p$, with
\[
	\norm{ x - x^{≤ m} }_p
	=
	∑_{n = m + 1}^∞ \abs{x_n}^p \,.
\]
It follows from the convergence of the series~$∑_{n = 0}^∞ \abs{x_n}^p$ that~$∑_{n = m + 1}^∞ \abs{x_n}^p \to 0$ as~$m \to ∞$.
There hence exists some sufficiently large index~$m$ for which
\[
	\norm{ x - x^{≤ m} }_p < \frac{ε}{2} \,.
\]
We now approximate the real numbers~$x_0, \dotsc, x_m$ by rational numbers~$y_0, \dotsc, y_m$ such that
\[
	\abs{x_i - y_i}^p < \frac{ε}{2 (m + 1)}
\]
for every~$i = 0, \dotsc, n$.
By setting~$y_n = 0$ for every~$n > m$ we get a sequence~$y$ that is an element of~$D$, and which satisfies~$\norm{x^{≤ m} - y}_p < ε / 2$.
It follows that
\[
	\norm{x - y}_p
	≤
	\norm{x - x^{≤ m}}_p + \norm{x^{≤ m} - y}_p
	<
	\frac{ε}{2} + \frac{ε}{2}
	=
	ε \,.
\]
According to \cref{characterization of dense subsets of metric spaces}, we have shown that~$D$ is dense in~$ℓ_p$.
As~$D$ is countable, this shows that~$ℓ_p$ is separable.

We now show that~$ℓ_∞$ is not separable.
For every subset~$A$ of~$ℕ$ we consider its characteristic function~$χ^A$ as a sequence, given by
\[
	χ^A_n
	=
	\begin{cases*}
		1 & if~$n ∈ A$, \\
		0 & otherwise.
	\end{cases*}
\]
For every two distinct subsets~$A$ and~$B$ of~$ℕ$ there exists an element~$n$ of~$ℕ$ with either~$n ∈ A$ and~$n ∉ B$ or~$n ∉ A$ and~$n ∈ B$.
We have in either case~$\abs{χ^A_n - χ^B_n} = 1$, and thus overall
\[
	\norm[\big]{χ^A - χ^B}_∞ = 1 \,.
\]
The set
\[
	P = \{ χ^A \suchthat A ⊆ ℕ \}
\]
is uncountable, and we have seen that for every two distinct elements~$x$ and~$y$ of~$P$ we have~$\norm{x - y}_∞ = 1$.
According to \cref{criterion for non-separable in metric spaces}, the existence of this set~$P$ shows that~$ℓ_∞$ is not separable.

