\subsection{The Functor~\texorpdfstring{$β$}{β}}

We will verify the claimed functorialities and naturalities.


\subsubsection{Functoriality of~$β$}

\begin{proposition}
	\label{functoriality of pushforward of filters}
	Let~$X$,~$Y$ and~$Z$ be sets and let~$f \colon X \to Y$ and~$g \colon Y \to Z$ be two maps.
	\begin{enumerate}

		\item
			We have~$(\id_X)_* \filter{F} = \filter{F}$ for every filter~$\filter{F}$ on~$X$.

		\item
			We have~$g_* f_* \filter{F} = (g f)_* \filter{F}$ for every filter~$\filter{F}$ on~$X$.

	\end{enumerate}
\end{proposition}

\begin{proof}
	\leavevmode
	\begin{enumerate}

		\item
			We have for every subset~$A$ of~$X$ the chain of equivalences
			\[
				A ∈ (\id_X)_* \filter{F}
				\iff
				\id_X^{-1} A ∈ \filter{F}
				\iff
				A ∈ \filter{F} \,,
			\]
			and therefore the equality~$(\id_X)_* \filter{F} = \filter{F}$.

		\item
			We have for every subset~$A$ of~$Z$ the chain of equivalences
			\begin{align*}
				A ∈ g_* f_* \filter{F}
				\iff
				g^{-1} A ∈ f_* \filter{F}
				\iff
				f^{-1} g^{-1} A ∈ \filter{F}
				&\iff
				(g f)^{-1} A ∈ \filter{F} \\
				&\iff
				A ∈ (g f)_* \filter{F} \,,
			\end{align*}
			and therefore the equality~$g_* f_* \filter{F} = (g f)_* \filter{F}$.
		\qedhere

	\end{enumerate}
\end{proof}

It follows from the above~\lcnamecref{functoriality of pushforward of filters} and Theorem~3.13 that the assignment~$β \colon \Set \to \Set$ is indeed a functor.



\subsubsection{Naturality of~$α$}

\begin{lemma}
	\label{pushforward of neighbourhood filter}
	Let~$(X, \top{T}^X)$ and~$(Y, \top{T}^Y)$ be two topological spaces and let~$f$ be a continuous map from~$(X, \top{T}^X)$ to~$(Y, \top{T}^Y)$.
	For every point~$x$ in~$X$, we have
	\[
		\top{T}^Y_{f x} ⊆ f_* \top{T}^X_x \,.
	\]
\end{lemma}

\begin{proof}
	Let~$U$ be an open neighbourhood of~$f x$.
	The preimage~$f^{-1} U$ is an open neighbourhood of~$x$ because~$f$ is continuous.
	Therefore,~$f^{-1} U$ belongs to~$\top{T}^X_x$, which means that~$U$ belongs to~$\top{T}^Y_{f x}$.
\end{proof}

\begin{proposition}
	\label{naturality of limit of filters}
	Let~$X$ be a topological space and let~$\filter{F}$ be a filter on~$X$ converging to some point~$x$ in~$X$.
	Let~$Y$ be another topological space and let~$f$ be a continuous map from~$X$ to~$Y$.
	Then the filter~$f_* \filter{F}$ converges to~$f x$.
\end{proposition}

\begin{proof}
	Let~$\top{T}^X$ and~$\top{T}^Y$ denote the topologies on~$X$ and~$Y$ respectively.
	That~$\filter{F}$ converges to~$x$ means that~$\top{T}^X_x ⊆ \filter{F}$.
	It then follows that
	\[
		\top{T}^Y_{f x}
		⊆
		f_* \top{T}^X_x
		⊆
		f_* \filter{F} \,,
	\]
	where the first inclusion follows from~\cref{pushforward of neighbourhood filter}.
	The inclusion~$\top{T}^Y_{f x} ⊆ f_* \filter{F}$ tells us that the filter~$f_* \filter{F}$ converges to~$f x$.
\end{proof}

\Cref{naturality of limit of filters} tells us that the assignment~$α_X \colon β X \to X$, where~$X$ is a compact Hausdorff space, is natural.
The map~$α_X$ is also bijective for every compact Hausdorff space~$X$.
We have therefore constructed a natural isomorphism between the functor~$β \colon \CptHaus \to \Set$ and the forgetful functor from~$\CptHaus$ to~$\Set$.



\subsubsection{Naturality of~$η$}

\begin{proposition}
	\label{pushforward of principal filter}
	Let~$X$ and~$Y$ be two sets and let~$f$ be a map from~$X$ to~$Y$.
	For every point~$x$ in~$X$, we have~$f_* P_x = P_{f x}$.
\end{proposition}

\begin{proof}[First proof]
	We have for every subset~$A$ of~$Y$ the chain of equivalences
	\[
		A ∈ f_* P_x
		\iff
		f^{-1} A ∈ P_x
		\iff
		x ∈ f^{-1} A
		\iff
		f x ∈ A
		\iff
		A ∈ P_{f x} \,,
	\]
	and therefore the equality~$f_* P_x = P_{f x}$.
\end{proof}

\begin{proof}[Second proof]
	Let~$\top{T}^X$ and~$\top{T}^Y$ denote the discrete topologies on~$X$ and~$Y$ respectively.
	The map~$f$ is continuous with respect to the discrete topologies, and we have~$\top{T}^X_x = P_x$ and~$\top{T}^Y_{f x} = P_{f x}$.
	It follows from \cref{pushforward of neighbourhood filter} that
	\[
		P_{f x}
		=
		\top{T}^Y_{f x}
		⊆
		f_* \top{T}^X_x
		=
		f_* P_x \,.
	\]
	We know from Theorem~3.13 that the pushforward~$f_* P_x$ is again an ultrafilter.
	So~$P_{f x}$ and~$f_* P_x$ are ultrafilters with~$P_{f x} ⊆ f_* P_x$, whence already~$P_{f x} = f_* P_x$.
\end{proof}

The above \lcnamecref{pushforward of principal filter} tells us that the transformation~$η \colon \id_{\Set} \to β$ given by~$η_X x = P_x$ is natural.
